---
title: "Project DATA MINING"
author: "DEEP SAGAR KARKI"
date: "3/26/2022"
output: pdf_document
---

---
title: "Project"
author: "DSK"
date: "3/10/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Packages
```{r message=FALSE}
library(dplyr)
```


### Data
```{r}
library(magrittr)
library(tidyverse)
data <- read.csv("C:/Users/14193/Downloads/Life Expectancy Data.csv") %>%
         filter(Year==2014) 
dat=data[,-c(1,2)]
dat$Status=as.factor(dat$Status)
levels(dat$Status)=c("no","yes")
```

### Partitioning The Data
```{r}
RNGkind(sample.kind = "Rounding")
set.seed(0)

partition.2 <- function(data, prop.train){
  # select a random sample of size = prop.train % of total records
  selected <- sample(1:nrow(data), round(nrow(data)*prop.train), replace = FALSE) 
  # create training data which has prop.train % of total records
  data.train <- data[selected,]
  # create validation data
  rest <- setdiff(1:nrow(data), selected)
  data.test <- data[rest,]
  return(list(data.train=data.train, data.test=data.test))
}

partitioned=partition.2(dat,0.7)
training.data=partitioned$data.train
test.data=partitioned$data.test


```



### Imputation on training data and using the attributes of training data on test data 


```{r}

sapply(training.data, anyNA)

# BMI,Total.expenditure,GDP,thinness.5.9,Hepatities,Population,Income composition,thinnes.1.19,schooling are the variables that were found to have missing values in training data




sapply(test.data, anyNA)
# HepatitisB,BMI, TOTAL EXPENDITURE,GDP,thinness.5.9,Population,Income composition,Alcohol,thinnnes.1.19,schooling are the variables with missing values in the test data set. 

# We will first replace the missing values with the median values of the respective columns.


med.BMI.train <- median(training.data$BMI, na.rm = TRUE)
training.data$BMI[is.na(training.data$BMI)] <- med.BMI.train
test.data$BMI[is.na(test.data$BMI)] <- med.BMI.train


med.Hepatitis.train <- median(training.data$Hepatitis.B, na.rm = TRUE)
training.data$Hepatitis.B[is.na(training.data$Hepatitis.B)] <- med.Hepatitis.train
test.data$Hepatitis.B[is.na(test.data$Hepatitis.B)] <- med.Hepatitis.train

med.total_expenditure.train <- median(training.data$Total.expenditure, na.rm = TRUE)
training.data$Total.expenditure[is.na(training.data$Total.expenditure)] <- med.total_expenditure.train
test.data$Total.expenditure[is.na(test.data$Total.expenditure)] <- med.total_expenditure.train


med.GDP.train <- median(training.data$GDP, na.rm = TRUE)
training.data$GDP[is.na(training.data$GDP)] <- med.GDP.train
test.data$GDP[is.na(test.data$GDP)] <- med.GDP.train

med.thinnes1_19.train <- median(training.data$thinness..1.19.years, na.rm = TRUE)
training.data$thinness..1.19.years[is.na(training.data$thinness..1.19.years)] <- med.thinnes1_19.train
test.data$thinness..1.19.years[is.na(test.data$thinness..1.19.years)] <-  med.thinnes1_19.train


med.population.train <- median(training.data$Population, na.rm = TRUE)
training.data$Population[is.na(training.data$Population)] <- med.population.train
test.data$Population[is.na(test.data$Population)] <- med.population.train


med.income.train <- median(training.data$Income.composition.of.resources, na.rm = TRUE)
training.data$Income.composition.of.resources[is.na(training.data$Income.composition.of.resources)] <- med.income.train
test.data$Income.composition.of.resources[is.na(test.data$Income.composition.of.resources)] <- med.income.train

med.under5_deaths.train <- median(training.data$under.five.deaths, na.rm = TRUE)
training.data$under.five.deaths[is.na(training.data$under.five.deaths)] <- med.under5_deaths.train
test.data$under.five.deaths[is.na(test.data$under.five.deaths)] <- med.under5_deaths.train


med.alcohol.train <- median(training.data$Alcohol, na.rm = TRUE)
training.data$Alcohol[is.na(training.data$Alcohol)] <- med.alcohol.train
test.data$Alcohol[is.na(test.data$Alcohol)] <- med.alcohol.train



med.thinnes5_9.train <- median(training.data$thinness.5.9.years, na.rm = TRUE)
training.data$thinness.5.9.years[is.na(training.data$thinness.5.9.years)] <- med.thinnes5_9.train
test.data$thinness.5.9.years[is.na(test.data$thinness.5.9.years)] <- med.thinnes5_9.train 


med.schooling.train <- median(training.data$Schooling, na.rm = TRUE)
training.data$Schooling[is.na(training.data$Schooling)] <- med.schooling.train
test.data$Schooling[is.na(test.data$Schooling)] <- med.schooling.train

```


### Fitting a model on the training data
```{r}
full_lm <- lm(Life.expectancy~., data=training.data)
```




### Residual analysis


```{r}
## Check for linearity

plot(x=training.data$Status, y=full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Status")
abline(h=0)
plot(training.data$Adult.Mortality, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. adult mortality")
abline(h=0)

plot(training.data$infant.deaths, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. infant deaths")
abline(h=0)

plot(training.data$infant.deaths, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. alcohol")
abline(h=0)


plot(training.data$percentage.expenditure, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. percent expenditure")
abline(h=0)


plot(training.data$Hepatitis.B, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. hepatitis B")
abline(h=0)



plot(training.data$Measles, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Measles")
abline(h=0)

plot(training.data$BMI, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. BMI")
abline(h=0)

plot(training.data$under.five.deaths, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. under five deaths")
abline(h=0)


plot(training.data$Polio, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Polio")
abline(h=0)

plot(training.data$Total.expenditure, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. total expenditure")
abline(h=0)


plot(training.data$Diphtheria, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Diphtheria")
abline(h=0)

plot(training.data$HIV.AIDS, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. HIV/AIDS")
abline(h=0)

plot(training.data$GDP, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. GDP")
abline(h=0)

plot(training.data$Population, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. population")
abline(h=0)


plot(training.data$thinness..1.19.years, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. thiness.1.19 years")
abline(h=0)

plot(training.data$thinness.5.9.years, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. thiness.5.19 years")
abline(h=0)

plot(training.data$Income.composition.of.resources, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Income composition of resources")
abline(h=0)


plot(training.data$Schooling, full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Schooling")
abline(h=0)




## Check for zero mean and constant variance of random error
plot(full_lm$fitted.values, full_lm$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)

## Check for normality of random error
qqnorm(full_lm$residuals)
qqline(full_lm$residuals)



```


Definitely needs a transformation as only some of the linearity plots are satisfactory. We are okay with constant variance.




## Applying the transformations


```{r}

##### lets log transformation on response variable #####

new.training.data <- training.data
new.training.data$infant.deaths <- new.training.data$infant.deaths^(0.5)

new.training.data$Alcohol <- new.training.data$Alcohol^(0.5)
new.training.data$percentage.expenditure <- new.training.data$percentage.expenditure^(0.5)
new.training.data$Hepatitis.B <- new.training.data$Hepatitis.B^(0.5)

new.training.data$Measles <- new.training.data$Measles^(0.5)

new.training.data$under.five.deaths <- new.training.data$under.five.deaths^(0.5)

new.training.data$Polio <- new.training.data$Polio^(0.5)

new.training.data$Diphtheria<- new.training.data$Diphtheria^(0.5)

new.training.data$HIV.AIDS <- new.training.data$HIV.AIDS^(0.5)

new.training.data$GDP <- new.training.data$GDP^(0.5)

new.training.data$Population <- new.training.data$Population^(0.5)


new.training.data$Measles <- new.training.data$Measles^(0.5)

new.training.data$thinness..1.19.years <- new.training.data$thinness..1.19.years^(0.5)

new.training.data$thinness.5.9.years <- new.training.data$thinness.5.9.years^(0.5)

new_full_lm <- lm(Life.expectancy ~. , data = new.training.data)
summary(new_full_lm )



## Check for linearity
par(mfcol = c(2, 2))
plot(new.training.data$infant.deaths, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. infant deaths")
abline(h=0)


plot(new.training.data$Alcohol, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs.Alcohol")
abline(h=0)


plot(new.training.data$percentage.expenditure, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs.Percent expenditure")
abline(h=0)


plot(new.training.data$Hepatitis.B, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs.Hepatitis B")
abline(h=0)


plot(new.training.data$Measles, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs.measles")
abline(h=0)


plot(new.training.data$under.five.deaths, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. under 5 deaths")
abline(h=0)


plot(new.training.data$Polio, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Polio")
abline(h=0)


plot(new.training.data$Diphtheria, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Diphtheria")
abline(h=0)

plot(new.training.data$HIV.AIDS, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. HIV")
abline(h=0)

plot(new.training.data$GDP, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. GDP")
abline(h=0)

plot(new.training.data$Population, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. Population")
abline(h=0)

plot(new.training.data$thinness..1.19.years, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. thinness1.19")
abline(h=0)



plot(new.training.data$thinness..1.19.years, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. thinness1.19")
abline(h=0)

plot(new.training.data$thinness.5.9.years, new_full_lm$residuals, 
     main = "Check for linearity \n Residuals vs. thinness5.9")
abline(h=0)











# Check for zero mean and constant variance of random error
plot(new_full_lm$fitted.values, new_full_lm$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)

## Check for normality of random error
qqnorm(new_full_lm$residuals)
qqline(new_full_lm$residuals)
```

 So, the new training model is 
 
```{r}
summary(new_full_lm)
 
```

Adjusted R _squared is pretty high.Residual standard error is also low.


## We will work on this model.