---
title: "REGRESSION-CLUSTERED-DATA"
author: "DEEP SAGAR KARKI"
date: "4/13/2022"
output: pdf_document
---

```{r}
cluster_1=read.csv("~/Downloads/Data-Mining-Final-Project-main/Clustered Data/cluster1.csv")
cluster_2=read.csv("~/Downloads/Data-Mining-Final-Project-main/Clustered Data/cluster2.csv")
cluster_3=read.csv("~/Downloads/Data-Mining-Final-Project-main/Clustered Data/cluster3.csv")
cluster_4=read.csv("~/Downloads/Data-Mining-Final-Project-main/Clustered Data/cluster4.csv")

```



## Graph before trasnformation

```{r}

par(mfrow = c(3, 4))
plot(cluster_1$Diptheria,cluster_1$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot(cluster_1$BMI,cluster_1$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot((cluster_1$Alcohol),cluster_1$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

plot(cluster_1$Adult.Mortality,cluster_1$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot(cluster_1$HepB,cluster_1$HALE_Birth, main="Check for linearity \n Response vs HEPB")

plot((cluster_1$HIV),(cluster_1$HALE_Birth), main="Check for linearity \n Response  vs HIV")


plot(cluster_1$infant_mortality,cluster_1$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot(cluster_1$percent.expenditure,cluster_1$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot(cluster_1$Polio,cluster_1$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot(cluster_1$thinness.5.19.years,cluster_1$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot(cluster_1$total.expenditure,cluster_1$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")







```
Some problems with HepB and HIV in linear relationship. All other variable show some trend of linear relationship.


We can fit a full model on all the regressors which will serve as a baseline model and the necessary transformation can be made. below is a full model for cluster 1.


```{r}
library(car)
lm_original_1=lm(HALE_Birth~., data=cluster_1[,-1])
vif(lm_original_1)


# The model suffers from huge multcollinearity


# CHECKING CONSTANTA VARIANCE ASSUMPTION
plot(lm_original_1$fitted.values, lm_original_1$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)


# The assumption for 0 meand and constant variance is satisfactory
```





## Linearity analysis after transformation

```{r}

par(mfrow = c(4, 4))
plot((cluster_1$Diptheria),cluster_1$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot((cluster_1$BMI),cluster_1$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot(log(cluster_1$Alcohol+0.5),cluster_1$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

plot((cluster_1$Adult.Mortality),cluster_1$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot((cluster_1$HepB)^5,cluster_1$HALE_Birth, main="Check for linearity \n Response vs HEPB")

plot((cluster_1$HIV),(cluster_1$HALE_Birth), main="Check for linearity \n Response  vs HIV")

## ignore this HIV

plot(log(cluster_1$infant_mortality),cluster_1$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot(log(cluster_1$percent.expenditure),cluster_1$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot((cluster_1$Polio),cluster_1$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_1$thinness.5.19.years),cluster_1$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot((cluster_1$total.expenditure),cluster_1$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# will not include HIV for linear regression

```


## Transformed clustered dataset
```{r}

cluster_1$logAlcohol=log(cluster_1$Alcohol+0.5)

cluster_1$HepB_power_5=cluster_1$HepB^5
cluster_1$logInfant=log(cluster_1$infant_mortality)

cluster_1$logpercent_expenditure=log(cluster_1$percent.expenditure)




```




Now we will fit the regression model first and see if the model suffers from multicollinearity:

```{r}

lm_cluster_1=lm(HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ thinness.5.19.years+total.expenditure , data = cluster_1)
summary(lm_cluster_1)
vif(lm_cluster_1)


```




Suffers from multicollinearity as VIF values of 19.35 for diptheria, 20.7629 for Polio and 4.20 for Adult Mortality indicate multicollinearity.





## Variable selection : Lasso regression

```{r}
#########################
### Lasso regression ####
#########################
library(caret)
set.seed(0)
RNGkind(sample.kind = "Rounding")
train_control <- trainControl(method="LOOCV")
glmnet.lasso <- train( HALE_Birth ~ Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ thinness.5.19.years+total.expenditure , data = cluster_1[,-1], method = "glmnet",
                      trControl = train_control, 
                      tuneGrid = expand.grid(alpha = 1,lambda = seq(0.001,0.1,by = 0.001)))

glmnet.lasso 
plot(glmnet.lasso)
# best coefficient
lasso.model <- coef(glmnet.lasso$finalModel, glmnet.lasso$bestTune$lambda)
lasso.model


```

Through lasso regression technique, Diphtheria was shrunk to 0. So, we fit the model without Diphtheria and look at multicollinearity again.


```{r}
lm_cluster_1_nodiph=lm(HALE_Birth ~Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ thinness.5.19.years+total.expenditure , data = cluster_1)
summary(lm_cluster_1_nodiph)
vif(lm_cluster_1_nodiph)

```
The model still suffers from multicollinearity. Now, Adult Mortality has vif values of 4.18 but this regressor has significance to the model based on the corresponding p-value. So, we drop the variable with the least significant in the model, which is logAlcohol.



## We repeat the process with cluster 2


## Graph before trasnformation

```{r}

par(mfrow = c(3, 4))
plot(cluster_2$Diptheria,cluster_2$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot(cluster_2$BMI,cluster_2$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot((cluster_2$Alcohol),cluster_2$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

# ALcohol needs log transformation

plot(cluster_2$Adult.Mortality,cluster_2$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot(cluster_2$HepB,cluster_2$HALE_Birth, main="Check for linearity \n Response vs HEPB")

# power to the 5th transformation

plot((cluster_2$HIV),(cluster_2$HALE_Birth), main="Check for linearity \n Response  vs HIV")

# Transformation with HIV is difficult. We ignore this one.


plot((cluster_2$infant_mortality),cluster_2$HALE_Birth, main="Check for linearity \n Response vs infant mortality")

# log transformation to infant mortality

plot((cluster_2$percent.expenditure),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")
# log transformation to percent expenditure

plot((cluster_2$Polio),cluster_2$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_2$thinness.5.19.years),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot((cluster_2$total.expenditure),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")


#We can fit a full model on all the regressors which will serve as a baseline model and the necessary transformation can be made

lm_original_2=lm(HALE_Birth~.,data=cluster_2[,-1])
vif(lm_original_2)


# CHECKING CONSTANTA VARIANCE ASSUMPTION
plot(lm_original_2$fitted.values, lm_original_2$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)


# The assumption for 0 meand and constant variance is satisfactory


```
The original model for cluster 2 suffers from high multicollinearity.




## Residual analysis. after transformation

```{r}

par(mfrow = c(4, 4))
plot((cluster_2$Diptheria),cluster_2$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot((cluster_2$BMI),cluster_2$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot(log(cluster_2$Alcohol+0.5),cluster_2$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

plot((cluster_2$Adult.Mortality),cluster_2$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot((cluster_2$HepB)^5,cluster_2$HALE_Birth, main="Check for linearity \n Response vs HEPB")

plot((cluster_2$HIV),(cluster_2$HALE_Birth), main="Check for linearity \n Response  vs HIV")

## ignore this HIV

plot(log(cluster_2$infant_mortality),cluster_2$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot(log(cluster_2$percent.expenditure),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot((cluster_2$Polio),cluster_2$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_2$thinness.5.19.years)^(1/2),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot((cluster_2$total.expenditure)^(1/2),cluster_2$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# will not include HIV for linear regression

```
## Transformed clustered dataset fro cluster 2
```{r}

cluster_2$logAlcohol=log10(cluster_2$Alcohol+0.5)


cluster_2$HepB_power_5=cluster_2$HepB^5
cluster_2$logInfant=log(cluster_2$infant_mortality)

cluster_2$logpercent_expenditure=log(cluster_2$percent.expenditure)

cluster_2$sqaurerootthinnes=(cluster_2$thinness.5.19.years)^(1/2)
cluster_2$squareroottotal_expenditure=(cluster_2$total.expenditure)^(1/2)

```

# had to do log(al+0.5), because after trasnformation we were getting -inf values




Now we will fit the regression model first and see if the model suffers from multicollinearity:

```{r}


lm_cluster_2=lm(HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ sqaurerootthinnes+squareroottotal_expenditure , data = cluster_2)
summary(lm_cluster_2)
vif(lm_cluster_2)


```

The model suffers from multicollinearity because Diphtheria 






## Variable selection : Lasso regression

```{r}
#########################
### Lasso regression ####
#########################

set.seed(0)
RNGkind(sample.kind = "Rounding")
train_control <- trainControl(method="LOOCV")
glmnet.lasso <- train( HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ sqaurerootthinnes+squareroottotal_expenditure , data = cluster_2[,-1], method = "glmnet",
                      trControl = train_control, 
                      tuneGrid = expand.grid(alpha = 1,lambda = seq(0.001,0.1,by = 0.001)))

glmnet.lasso 
plot(glmnet.lasso)
# best coefficient
lasso.model <- coef(glmnet.lasso$finalModel, glmnet.lasso$bestTune$lambda)
lasso.model


```

Diphtheria, BMI, Hepb_power5, have shrunked to 0. We will fit the model now.



```{r}
lm_cluster_2_nodip.BMI.HepB=lm(HALE_Birth ~Adult.Mortality+ logAlcohol  +Polio+logpercent_expenditure+logInfant + sqaurerootthinnes+squareroottotal_expenditure , data = cluster_2)
summary(lm_cluster_2_nodip.BMI.HepB)
vif(lm_cluster_2_nodip.BMI.HepB)

```
The model does not suffer from multicollinearity.











## We repeat the process with cluster 3


## Graph before trasnformation

```{r}

par(mfrow = c(3, 4))
plot(cluster_3$Diptheria,cluster_3$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot(cluster_3$BMI,cluster_3$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot((cluster_3$Alcohol),cluster_3$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

# ALcohol needs log transformation

plot(cluster_3$Adult.Mortality,cluster_3$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot(cluster_3$HepB,cluster_3$HALE_Birth, main="Check for linearity \n Response vs HEPB")

# power to the 5th transformation

plot((cluster_3$HIV),(cluster_3$HALE_Birth), main="Check for linearity \n Response  vs HIV")

# HIV needs log transformation


plot((cluster_3$infant_mortality),cluster_3$HALE_Birth, main="Check for linearity \n Response vs infant mortality")

# log transformation to infant mortality

plot((cluster_3$percent.expenditure),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")
# log transformation to percent expenditure

plot((cluster_3$Polio),cluster_3$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_3$thinness.5.19.years),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot((cluster_3$total.expenditure),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# total expenditure needs log transformation

#We can fit a full model on all the regressors which will serve as a baseline model and the necessary transformation can be made

lm_original_3=lm(HALE_Birth~.,data=cluster_3[,-1])
vif(lm_original_3)

# CHECKING CONSTANTA VARIANCE ASSUMPTION
plot(lm_original_3$fitted.values, lm_original_3$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)


# The assumption for 0 meand and constant variance is satisfactory



```
The original model for cluster 3 suffers from high multicollinearity..




## Linearity analysis  after transformation

```{r}

par(mfrow = c(4, 4))
plot((cluster_3$Diptheria),cluster_3$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot((cluster_3$BMI),cluster_3$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot(log(cluster_3$Alcohol+0.5),cluster_3$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

plot((cluster_3$Adult.Mortality),cluster_3$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot((cluster_3$HepB)^5,cluster_3$HALE_Birth, main="Check for linearity \n Response vs HEPB")

plot((cluster_3$HIV),(cluster_3$HALE_Birth), main="Check for linearity \n Response  vs HIV")



plot(log(cluster_3$infant_mortality),cluster_3$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot(log(cluster_3$percent.expenditure),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot((cluster_3$Polio),cluster_3$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_3$thinness.5.19.years),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot(log(cluster_3$total.expenditure),cluster_3$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# will not include HIV for linear regression

```
## Transformed clustered dataset fro cluster 2
```{r}

cluster_3$logAlcohol=log(cluster_3$Alcohol+0.5)


cluster_3$HepB_power_5=cluster_3$HepB^5
cluster_3$logInfant=log(cluster_3$infant_mortality)

cluster_3$logpercent_expenditure=log(cluster_3$percent.expenditure)

cluster_3$logtotal_expenditure=log(cluster_3$total.expenditure)


```

# had to do log(al+0.5), because after trasnformation we were getting -inf values




Now we will fit the regression model first and see if the model suffers from multicollinearity:

```{r}


lm_cluster_3=lm(HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ thinness.5.19.years+logtotal_expenditure , data = cluster_3)
summary(lm_cluster_3)
vif(lm_cluster_3)


```

The model suffers from multicollinearity






## Variable selection : Lasso regression

```{r}
#########################
### Lasso regression ####
#########################

set.seed(0)
RNGkind(sample.kind = "Rounding")
train_control <- trainControl(method="LOOCV")
glmnet.lasso <- train( HALE_Birth ~ Diptheria+Adult.Mortality+BMI + logAlcohol + HepB_power_5 +Polio+logpercent_expenditure+logInfant+ thinness.5.19.years+logtotal_expenditure , data = cluster_3[,-1], method = "glmnet",
                      trControl = train_control, 
                      tuneGrid = expand.grid(alpha = 1,lambda = seq(0.001,0.1,by = 0.001)))

glmnet.lasso 
plot(glmnet.lasso)
# best coefficient
lasso.model <- coef(glmnet.lasso$finalModel, glmnet.lasso$bestTune$lambda)
lasso.model


```

BMI, Polio,logpercent_expenditure and thinnes have shrunked to 0. We will fit the model now.



```{r}
lm_cluster_3_noBMI.polio.logpercent.thin=lm(HALE_Birth ~Diptheria+Adult.Mortality+logAlcohol+HepB_power_5+logInfant +total.expenditure , data = cluster_3)
summary(lm_cluster_3_noBMI.polio.logpercent.thin)
vif(lm_cluster_3_noBMI.polio.logpercent.thin)

```
The model suffers from multicollinearity. Now, since, logAlcohol  has least significance in the model, we will drip the variable and fit


## We repeat the process with cluster 4


## Graph before trasnformation

```{r}
library(car)
par(mfrow = c(3, 4))
plot(cluster_4$Diptheria,cluster_4$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot(cluster_4$BMI,cluster_4$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot((cluster_4$Alcohol),cluster_4$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

# ALcohol needs log transformation

plot(cluster_4$Adult.Mortality,cluster_4$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot((cluster_4$HepB),cluster_4$HALE_Birth, main="Check for linearity \n Response vs HEPB")

# power to the 5th transformation

plot((cluster_4$HIV),(cluster_4$HALE_Birth), main="Check for linearity \n Response  vs HIV")




plot((cluster_4$infant_mortality),cluster_4$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot((cluster_4$percent.expenditure),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot((cluster_4$Polio),cluster_4$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_4$thinness.5.19.years),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot((cluster_4$total.expenditure),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# total expenditure needs log transformation

#We can fit a full model on all the regressors which will serve as a baseline model and the necessary transformation can be made

lm_original_4=lm(HALE_Birth~.,data=cluster_4[,-1])
vif(lm_original_4)




# CHECKING CONSTANTA VARIANCE ASSUMPTION
plot(lm_original_4$fitted.values, lm_original_4$residuals, 
     main = "Check for 0 mean and constant var \n  Residual vs. fitted value")
abline(h=0)


# The assumption for 0 meand and constant variance is satisfactory



```
The original model for cluster 4 suffers from high multicollinearity..




## Residual analysis. after transformation

```{r}

par(mfrow = c(4, 4))
plot((cluster_4$Diptheria),cluster_4$HALE_Birth, main="Check for linearity \n Response vs Diphtheria")

plot((cluster_4$BMI),cluster_4$HALE_Birth, main="Check for linearity \n Response vs BMI")

plot(log(cluster_4$Alcohol+0.5),cluster_4$HALE_Birth, main="Check for linearity \n Response vs ALCOHOL")

plot((cluster_4$Adult.Mortality),cluster_4$HALE_Birth, main="Check for linearity \n Response vs ADULT MORTALITY")

plot((cluster_4$HepB)^5,cluster_4$HALE_Birth, main="Check for linearity \n Response vs HEPB")

plot((cluster_4$HIV),(cluster_4$HALE_Birth), main="Check for linearity \n Response  vs HIV")



plot((cluster_4$infant_mortality),cluster_4$HALE_Birth, main="Check for linearity \n Response vs infant mortality")


plot((cluster_4$percent.expenditure),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs percent expenditure")


plot((cluster_4$Polio),cluster_4$HALE_Birth, main="Check for linearity \n Response vs Polio")

plot((cluster_4$thinness.5.19.years),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs tthiness.5.19 ")

plot(log(cluster_4$total.expenditure),cluster_4$HALE_Birth, main="Check for linearity \n Response  vs total expenditure")

# will not include HIV for linear regression

```
## Transformed clustered dataset fro cluster 4
```{r}

cluster_4$logAlcohol=log(cluster_4$Alcohol+0.5)


cluster_4$HepB_power_5=cluster_4$HepB^5


cluster_4$logtotalexpenditure=log(cluster_4$total.expenditure)




```






Now we will fit the regression model first and see if the model suffers from multicollinearity:

```{r}


lm_cluster_4=lm(HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol +HIV+ HepB_power_5 +Polio+percent.expenditure+infant_mortality+ thinness.5.19.years+logtotalexpenditure , data = cluster_4)
summary(lm_cluster_4)
vif(lm_cluster_4)


```

The model suffers from multicollinearity










## Variable selection : Lasso regression

```{r}
#########################
### Lasso regression ####
#########################

set.seed(0)
RNGkind(sample.kind = "Rounding")
train_control <- trainControl(method="LOOCV")
glmnet.lasso <- train( HALE_Birth ~Diptheria+Adult.Mortality+BMI + logAlcohol + HIV+HepB_power_5 +Polio+percent.expenditure+infant_mortality+ thinness.5.19.years+logtotalexpenditure , data = cluster_4[,-1], method = "glmnet",
                      trControl = train_control, 
                      tuneGrid = expand.grid(alpha = 1,lambda = seq(0.001,0.1,by = 0.001)))

glmnet.lasso 
plot(glmnet.lasso)
# best coefficient
lasso.model <- coef(glmnet.lasso$finalModel, glmnet.lasso$bestTune$lambda)
lasso.model


```

Diphtheria, LogAlcohol,Polio, thinness and logtotalexpenditure were dropped.



Now we will fit the regression model 

```{r}

lm_cluster_4_dropped.lasso=lm( HALE_Birth ~Adult.Mortality+BMI + HIV+HepB_power_5 +percent.expenditure+infant_mortality,data=cluster_4)
summary(lm_cluster_4_dropped.lasso)
vif(lm_cluster_4_dropped.lasso)



```


The model again suffers from multicollinearity





`