---
title: "Austin's Clustering"
author: "L. Austin Hadamuscin"
date: "4/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r message=FALSE, warning=FALSE}
library(sampling)
library(tidyverse)
library(rworldmap)
library(mclust)
library(gtable)
```

## Data
```{r}
dat <- read.csv("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/imputedData.csv")
```

## Data Scaling
```{r}
 dat_scaled <- scale(select(dat, -Country, -HALE_Birth))
```

# k-means

```{r warning=FALSE}
RNGkind (sample.kind = "Rounding")
set.seed(0)

nseeds <- 1000
nk <- 20
 
seeds <- ceiling(runif(nseeds,0,900000))

# kmean_matrix  <- matrix(NA, nrow = nseeds, ncol = nk)
# 
# for (k in 1:nk) {
#   seed_iter = 0
#   for (s in seeds) {
# 
#     set.seed(s)
#     seed_iter <- seed_iter + 1
# 
#     kmean_matrix[seed_iter,k] <- kmeans(dat_scaled, centers = k)$tot.withinss
#   }
# }
# save(kmean_matrix, file = "kmean_matrix.RData")

load("kmean_matrix.RData")
```

```{r echo=FALSE}
# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/kMeansWSS.svg",
# width = 7.291667, height = 4.500000)
plot(1:nk,apply(kmean_matrix, 2, min), 
     type = "o",
     main = "k-Means",
     ylab = "Within Sum of Squares",
     xlab = "Number of Centroids")
# dev.off()
```



## Finding the optimal seeds
```{r}
kmean_seed2 <- seeds[which(kmean_matrix[,2]==min(kmean_matrix[,2]))[1]]
kmean_seed4 <- seeds[which(kmean_matrix[,4]==min(kmean_matrix[,4]))[1]]
```

## Doing the 4-means
```{r}
set.seed(kmean_seed2)
means2 <- kmeans(dat_scaled, centers = 2)
set.seed(kmean_seed4)
means4 <- kmeans(dat_scaled, centers = 4)
```

## Rescaling the cluster centers
```{r}
means2.centers.rescaled <-  means2$centers
for (j in 1:11){
  means2.centers.rescaled[,j] <- attributes(dat_scaled)$`scaled:center`[j] +
    means2$centers[,j]*attributes(dat_scaled)$`scaled:scale`[j]
}

means4.centers.rescaled <-  means4$centers
for (j in 1:11){
  means4.centers.rescaled[,j] <- attributes(dat_scaled)$`scaled:center`[j] +
    means4$centers[,j]*attributes(dat_scaled)$`scaled:scale`[j]
}
```


## Rearranging cluster groups
```{r}
km4id <- ifelse(means4$cluster == 1, 2, 
                ifelse(means4$cluster == 2, 3,
                       ifelse(means4$cluster == 3, 1, 4)))
```

## Grouping data into clusters
```{r}
kclust2 <- cbind(select(dat, Country, HALE_Birth),clusters = means2$cluster,dat)
kclust4 <- cbind(select(dat, Country, HALE_Birth),clusters = as.factor(km4id),select(dat, -Country, -HALE_Birth)) %>% as.data.frame()
```

## 2 Cluster Graph
```{r message=FALSE}
kclust2_map <- joinCountryData2Map(kclust2, joinCode = "NAME",
                                  nameJoinColumn = "Country")

# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/kMeans2World.svg",
#     width = 7.291667, height = 4.500000)
par(mar=c(0,0,0,0))
mapCountryData(mapToPlot = kclust2_map,
               nameColumnToPlot="clusters",
               catMethod="categorical",
               colourPalette = c("#E41A1C", "#4DAF4A"),
               mapTitle = "",
               oceanCol = "light blue",
               missingCountryCol = "white",
               borderCol = "black",
               lwd = .25,
               addLegend = F)
legend(x = -180, y = -47.5, legend = c("1","2"),
       title = "category", fill = c("#E41A1C", "#4DAF4A"))
# dev.off()
```

## 4 Cluster Graph
```{r}
kclust4_map <- joinCountryData2Map(kclust4, joinCode = "NAME",
                                   nameJoinColumn = "Country")

# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/kMeans4World.svg",
#     width = 7.291667, height = 4.500000)

par(mar=c(0,0,0,0))
mapCountryData(mapToPlot = kclust4_map,
               nameColumnToPlot="clusters",
               catMethod="categorical",
               colourPalette = RColorBrewer::brewer.pal(4,'Set1'),
               mapTitle = "",
               oceanCol = "light blue",
               missingCountryCol = "white",
               borderCol = "black",
               lwd = .25,
               addLegend = F)
legend(x = -180, y = -26, legend = c("1","2","3","4"),
       title = "category", fill = RColorBrewer::brewer.pal(4,'Set1'))
# dev.off() 
```


```{r}
theme_set(theme_bw())
theme_update(
  text = element_text(size=12),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank()
)



means4$centers %>% unname  %>% t %>% data.frame(names = c("Diptheria","BMI","Alcohol","Adult.Mortality","HepB","HIV","infant_mortality","percent.expenditure","Polio","thinness.5.19.years","total.expenditure"),.) %>% pivot_longer(cols = -1, names_to = "line") -> means4centers

means4centers$line<- ifelse(means4centers$line == "X1", "2", 
                            ifelse(means4centers$line == "X2", "3",
                                   ifelse(means4centers$line == "X3", "1", "4")))

# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/kMeans4centers.svg",
#     width = 7.291667, height = 4.500000)

ggplot(data = means4centers, 
       mapping = aes(x = names, y = value, color = line)) + 
  geom_point() +
  geom_line(mapping = aes(group = line))+
  scale_color_manual(values = RColorBrewer::brewer.pal(4,'Set1')) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, vjust = 1.1))  + 
  labs(color = "Clusters", y = "Scaled Means", x="",
       title = "k-Means Clusters") 

# dev.off()
  
```


# Model Based

```{r message=FALSE}
# mb_rep <- 20
# mb_matrix <- matrix(NA, 14, mb_rep)
# 
# mb_dat <- select(dat, -Country, -HALE_Birth)
# 
# for (k in 1:mb_rep) {
#   mb_matrix[,k] <- Mclust(mb_dat, k)$BIC
# }
# 
# save(mb_matrix,file = "mb_matrix.RData")


load("mb_matrix.RData")
```


```{r}
# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/ModelBasedBIC.svg",
# width = 7.291667, height = 4.500000)
plot(x=1:20,
     y=apply(mb_matrix, 2, function(i) min(i,na.rm = T)),
     type = "o",
     xlab = "Number of Components",
     ylab = "Negative BIC")
# dev.off()
```
- 4 looks like the best K.

## Model-based clustering with 4 clusters
```{r}
# mb4 <- Mclust(mb_dat, 4, verbose = F)
# save(mb4, file = "mb4.RData")

load("mb4.RData")
```

## Rearranging the clusters
```{r}
mb4id <- ifelse(mb4$classification == 3, 1, 
                ifelse(mb4$classification == 2, 2,
                       ifelse(mb4$classification == 1, 4, 3)))
```


## Grouping data into clusters
```{r}
mbclust4 <- cbind(select(dat, Country, HALE_Birth),clusters = as.factor(mb4id),dat)
```


## Clust 4 Graph
```{r echo=FALSE, message=FALSE}
# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/ModelBased4.svg",
# width = 7.291667, height = 4.500000)
mbclust4_map <- joinCountryData2Map(mbclust4, joinCode = "NAME", nameJoinColumn = "Country")
par(mar=c(0,0,0,0))
mapCountryData(mapToPlot = mbclust4_map,
               nameColumnToPlot="clusters",
               catMethod="categorical",
               colourPalette = RColorBrewer::brewer.pal(4,'Set1'),
               mapTitle = "",
               oceanCol = "light blue",
               missingCountryCol = "white",
               borderCol = "black",
               lwd = .25,
               addLegend = F)
legend(x = -180, y = -26, legend = c("1","2","3","4"),
       title = "category", fill = RColorBrewer::brewer.pal(4,'Set1'))
# dev.off()
```

```{r}
theme_set(theme_bw())
theme_update(
  text = element_text(size=12),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank()
)

scaledmeans <- apply(mb4$parameters$mean,1,scale) %>% t

mb4mean <- scaledmeans %>% unname() %>%
  data.frame(names = c("Diptheria","BMI","Alcohol","Adult.Mortality","HepB","HIV","infant_mortality","percent.expenditure","Polio","thinness.5.19.years","total.expenditure"),.) %>%
  pivot_longer(cols = -1, names_to = "clusters", values_to = "means")

mb4mean$clusters <- ifelse(mb4mean$clusters == "X3", "1", 
                           ifelse(mb4mean$clusters == "X2", "2",
                                  ifelse(mb4mean$clusters == "X1", "4", "3")))

# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/mb4means.svg",
#     width = 7.291667, height = 4.500000)

ggplot(data = mb4mean, 
       mapping = aes(x = names, y = means, color = clusters)) + 
  geom_point() +
  geom_line(mapping = aes(group = clusters))+
  scale_color_manual(values = RColorBrewer::brewer.pal(4,'Set1')) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, vjust = 1.1))  + 
  labs(color = "Clusters", y = "Scaled Means", x="",
       title = "Model Based Clusters") 

# dev.off()
```


```{r}
mb4$parameters$mean %>% unname() %>%
  data.frame(names = c("Diptheria","BMI","Alcohol","Adult.Mortality","HepB","HIV","infant_mortality","percent.expenditure","Polio","thinness.5.19.years","total.expenditure"),.)
```



# Partition and save the data with the 4-means cluster.
```{r}
cluster1 <- kclust4 %>% filter(clusters == 1) %>% select(-clusters)
cluster2 <- kclust4 %>% filter(clusters == 2) %>% select(-clusters)
cluster3 <- kclust4 %>% filter(clusters == 3) %>% select(-clusters)
cluster4 <- kclust4 %>% filter(clusters == 4) %>% select(-clusters)

# write.csv(x = cluster1,
#           file = "C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Clustered Data/cluster1.csv",
#           row.names = F)
# write.csv(x = cluster2,
#           file = "C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Clustered Data/cluster2.csv",
#           row.names = F)
# write.csv(x = cluster3,
#           file = "C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Clustered Data/cluster3.csv",
#           row.names = F)
# write.csv(x = cluster4,
#           file = "C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Clustered Data/cluster4.csv",
#           row.names = F)
```





# Making graphs
```{r}
theme_set(theme_bw())
theme_update(
  text = element_text(size=12),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.background = element_blank()
)
ggplot(data = means4centers, 
       mapping = aes(x = names, y = value, color = line)) + 
  geom_point() +
  geom_line(mapping = aes(group = line)) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        plot.margin=grid::unit(c(0,0,-3,0), "mm"),
        panel.grid.major.x = element_line()) +
  scale_color_manual(values = RColorBrewer::brewer.pal(4,'Set1'))  + 
  labs(color = "Clusters", y = "k-Means Clusters", x="") -> p1

ggplot(data = mb4mean, 
       mapping = aes(x = names, y = means, color = clusters)) + 
  geom_point() +
  geom_line(mapping = aes(group = clusters)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(4,'Set1')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1.01,vjust = 1.05),
        plot.margin=grid::unit(c(-3,0,0,0), "mm"),
        panel.grid.major.x = element_line()) + 
  labs(color = "Clusters", y = "Model Based Clusters", x="")  -> p2


g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p2)

g <- rbind(g1, g2, size = "first")

# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/stackedClusterMeans.svg",
#     width = 6.5, height = 4.5)
grid.newpage()
grid.draw(g)
# dev.off()
```

```{r echo=FALSE, message=FALSE}
# svg("C:/Users/hadamul/OneDrive/Graduate school/Semester 4/Data Mining/Data-Mining-Final-Project/Plots/stackedMaps.svg",
# width = 6.5, height = 4.500000)

par(mar=c(0,0,0,0), mfrow = c(1,2))
mapCountryData(mapToPlot = kclust4_map,
               nameColumnToPlot="clusters",
               catMethod="categorical",
               colourPalette = RColorBrewer::brewer.pal(4,'Set1'),
               mapTitle = "",
               oceanCol = "light blue",
               missingCountryCol = "white",
               borderCol = "black",
               lwd = .25,
               addLegend = F,
               aspect = 1)
# legend(x = -180, y = -26, legend = c("1","2","3","4"),
#        title = "category", fill = RColorBrewer::brewer.pal(4,'Set1'))

mapCountryData(mapToPlot = mbclust4_map,
               nameColumnToPlot="clusters",
               catMethod="categorical",
               colourPalette = RColorBrewer::brewer.pal(4,'Set1'),
               mapTitle = "",
               oceanCol = "light blue",
               missingCountryCol = "white",
               borderCol = "black",
               lwd = .25,
               addLegend = F,)
# legend(x = -180, y = -26, legend = c("1","2","3","4"),
#        title = "category", fill = RColorBrewer::brewer.pal(4,'Set1'))
# dev.off()
```

